Uber Data Analysis Project

#1. Importing Libraries
```{r}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
```
```{r}
setwd("~/Documents/projects/uber")
```

```{r}
apr <- read.csv("uber-raw-data-apr14.csv")
may <- read.csv("uber-raw-data-may14.csv")
jun <- read.csv("uber-raw-data-jun14.csv")
jul <- read.csv("uber-raw-data-jul14.csv")
aug <- read.csv("uber-raw-data-aug14.csv")
sep <- read.csv("uber-raw-data-sep14.csv")
```

```{r}
data <- rbind(apr, may, jun, jul, aug, sep)
data
```
```{r}
data$Date.Time <- as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S")
data$Time <- format(as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
data$Date.Time <- ymd_hms(data$Date.Time)
data

```

```{r}
data$day <- factor(day(data$Date.Time))
data$month <- factor(month(data$Date.Time, label = TRUE))
data$year <- factor(year(data$Date.Time))
data$dayofweek <- factor(wday(data$Date.Time, label = TRUE))
data
```
```{r}
data$hour <- factor(hour(hms(data$Time)))
data$minute <- factor(minute(hms(data$Time)))
data$second <- factor(second(hms(data$Time)))
data
```

#2. Plotting the trips by the hours in a day
```{r}
summarize(group_by(data,hour),Total = n())
```

```{r}
hour_data <- dplyr::summarize(group_by(data,hour),Total = n())
datatable(hour_data)
```
```{r}
ggplot(hour_data, aes(hour, Total)) + 
        geom_bar( stat = "identity", fill = "steelblue", color = "red") +
           ggtitle("Trips Every Hour") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)

month_hour <- data %>%
          group_by(month, hour) %>%
             dplyr::summarize(Total = n())

ggplot(month_hour, aes(hour, Total, fill = month)) + 
       geom_bar( stat = "identity") +
          ggtitle("Trips by Hour and Month") +
           scale_y_continuous(labels = comma)
```
#3. Plotting data by trips during every day of the month
 We observe from the resulting visualization that 30th of the month had the highest trips in the year which is mostly contributed by the month of April.
```{r}
day_group <- data %>%
          group_by(day) %>%
             dplyr::summarize(Total = n()) 
datatable(day_group)
```

```{r}
ggplot(day_group, aes(day, Total)) + 
        geom_bar( stat = "identity", fill = "steelblue") +
           ggtitle("Trips Every Day") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)

day_month_group <- data %>%
         group_by(month, day) %>%
             dplyr::summarize(Total = n())

```
```{r}
colors = c("#CC1011", "#665555", "#05a399", "#cfcaca", "#f5e840", "#0683c9", "#e075b0")
```

```{r}
ggplot(day_month_group, aes(day, Total, fill = month)) + 
        geom_bar( stat = "identity") +
           ggtitle("Trips by Day and Month") +
            scale_y_continuous(labels = comma) +
            scale_fill_manual(values = colors)
```
#4. Number of Trips taking place during months in a year

```{r}
month_group <- data %>%
          group_by(month) %>%
             dplyr::summarize(Total = n()) 
datatable(month_group)
```
```{r}
ggplot(month_group , aes(month, Total, fill = month)) + 
        geom_bar( stat = "identity") +
           ggtitle("Trips by Month") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma) +
            scale_fill_manual(values = colors)
```
```{r}
month_weekday <- data %>%
         group_by(month, dayofweek) %>%
             dplyr::summarize(Total = n())

ggplot(month_weekday, aes(month, Total, fill = dayofweek)) + 
       geom_bar( stat = "identity", position = "dodge") +
          ggtitle("Trips by Day and Month") +
           scale_y_continuous(labels = comma) +
           scale_fill_manual(values = colors)
```

```{r}
ggplot(data, aes(Base)) + 
 geom_bar(fill = "darkred") +
 scale_y_continuous(labels = comma) +
 ggtitle("Trips by Bases")
```
```{r}
ggplot(data, aes(Base, fill = month)) + 
 geom_bar(position = "dodge") +
 scale_y_continuous(labels = comma) +
 ggtitle("Trips by Bases and Month") +
 scale_fill_manual(values = colors)
```
```{r}
ggplot(data, aes(Base, fill = dayofweek)) + 
 geom_bar(position = "dodge") +
 scale_y_continuous(labels = comma) +
 ggtitle("Trips by Bases and DayofWeek") +
 scale_fill_manual(values = colors)
```
#5. Creating a Heatmap visualization of day, hour and month

```{r}
day_and_hour <- data %>%
         group_by(day, hour) %>%
            dplyr::summarize(Total = n())

datatable(day_and_hour)
```
```{r}
ggplot(day_and_hour, aes(day, hour, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Hour and Day")
```

```{r}
ggplot(day_month_group, aes(day, month, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Month and Day")
```

```{r}
ggplot(month_weekday, aes(dayofweek, month, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Month and Day of Week")
```
```{r}
month_base <-  data %>%
                    group_by(Base, month) %>%
                        dplyr::summarize(Total = n()) 

day0fweek_bases <-  data %>%
                    group_by(Base, dayofweek) %>%
                        dplyr::summarize(Total = n()) 

ggplot(month_base, aes(Base, month, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Month and Bases")
```
```{r}
ggplot(day0fweek_bases, aes(Base, dayofweek, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Bases and Day of Week")

```
#6. Creating a map visualization of rides in New York

```{r}
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004

ggplot(data, aes(x=Lon, y=Lat)) +
  geom_point(size=1, color = "blue") +
     scale_x_continuous(limits=c(min_long, max_long)) +
      scale_y_continuous(limits=c(min_lat, max_lat)) +
        theme_map() +
           ggtitle("NYC MAP BASED ON UBER RIDES DURING 2014 (APR-SEP)")

ggplot(data, aes(x=Lon, y=Lat, color = Base)) +
  geom_point(size=1) +
     scale_x_continuous(limits=c(min_long, max_long)) +
      scale_y_continuous(limits=c(min_lat, max_lat)) +
       theme_map() +
          ggtitle("NYC MAP BASED ON UBER RIDES DURING 2014 (APR-SEP) by BASE")
```

